---
title: "Introduction to the dcce Package: DCCE Estimation for Panel Data"
author: "Mustapha Wasseja"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Introduction to the dcce Package: DCCE Estimation for Panel Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: references.bib
link-citations: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  fig.width = 7,
  fig.height = 4.5,
  out.width = "100%"
)
```

# Introduction

Macro-panel datasets---large panels of countries, regions, or industries observed over many
time periods---are increasingly central to empirical economic research. A pervasive
feature of such panels is **cross-sectional dependence (CSD)**: the residuals of unit $i$
are correlated with the residuals of unit $j$ even after controlling for observable
explanatory variables. Ignoring CSD causes standard panel estimators to produce
inconsistent estimates and badly sized tests.

The `dcce` package implements the family of **Common Correlated Effects (CCE)**
estimators introduced by @pesaran2006estimation and extended to dynamic panels by
@chudik2015common. These estimators filter out unobserved common factors---the
source of CSD---by augmenting each unit's regression with cross-sectional averages
(CSAs) of the dependent variable and the regressors. The package further implements
long-run estimators (CS-ARDL, CS-DL), an extensive CD test suite, the exponent of
cross-sectional dependence, and information criteria for CSA lag selection.

**Key features:**

- Mean Group (MG), CCE-MG, Dynamic CCE (DCCE), Regularized CCE (rCCE)
- Long-run estimators: CS-ARDL (@chudik2016long), CS-DL, Pooled Mean Group (PMG, @shin1999ardl)
- CD test suite: Pesaran CD (@pesaran2015testing), CDw (@juodis2022), PEA (@fan2015power), CD* (@pesaran2021cdstar)
- Exponent of cross-sectional dependence (@bailey2016exponent, @bailey2019exponent)
- Information criteria for CSA selection (@margaritella2023information)
- Rank condition classifier (@devos2024rank)
- Cross-section and wild bootstrap inference
- `broom`-compatible `tidy()` and `glance()` methods

```{r load-package}
library(dcce)
```

# The Problem: Cross-Sectional Dependence

## What is cross-sectional dependence?

In a panel with $N$ units and $T$ time periods, suppose the data generating process is:

$$
y_{it} = \alpha_i + \boldsymbol{\beta}_i' \mathbf{x}_{it} + \varepsilon_{it},
$$

where $\varepsilon_{it}$ is the error term. Standard panel estimators assume
$\mathbb{E}[\varepsilon_{it}\varepsilon_{jt}] = 0$ for $i \neq j$. In macro
panels, this assumption is routinely violated because units share common shocks
(global business cycles, oil price shocks, financial crises) and are connected via
trade, capital flows, and technology spillovers.

## Consequences of ignoring CSD

When the errors contain unobserved common factors,
$$
\varepsilon_{it} = \boldsymbol{\lambda}_i' \mathbf{f}_t + u_{it},
$$
standard fixed-effects and first-differencing estimators are inconsistent if the
regressors are also driven by $\mathbf{f}_t$---a situation called *strong* CSD. In
that case, $\mathbf{x}_{it}$ and $\varepsilon_{it}$ share a common component
($\boldsymbol{\lambda}_i' \mathbf{f}_t$), inducing endogeneity. @pesaran2006estimation
shows the resulting bias can be substantial.

## Testing for CSD: the `pcd_test()` function

Before estimating, we should test for CSD. The Pesaran [-@pesaran2015testing] CD
statistic is the benchmark:

$$
\text{CD} = \sqrt{\frac{2T}{N(N-1)}} \sum_{i=1}^{N-1} \sum_{j=i+1}^{N}
             \sqrt{T_{ij}}\, \hat{\rho}_{ij} \xrightarrow{d} \mathcal{N}(0,1),
$$

where $\hat{\rho}_{ij}$ is the sample correlation of the residuals of units $i$
and $j$. Under the null of no CSD, $\text{CD} \to \mathcal{N}(0,1)$ as
$N, T \to \infty$.

We illustrate with the `pwt8` dataset: 93 countries, 1960--2007, adapted from
@ditzen2018estimating.

```{r pwt8-load}
data(pwt8, package = "dcce")
str(pwt8)
```

Run a naive pooled OLS of GDP growth on factor inputs, ignoring CSD:

```{r cd-ols}
# Pooled OLS on complete cases
pwt_cc <- na.omit(pwt8[, c("country", "year", "d_log_rgdpo",
                             "log_hc", "log_ck", "log_ngd")])
fit_ols <- lm(d_log_rgdpo ~ log_hc + log_ck + log_ngd, data = pwt_cc)

# CD test on OLS residuals
cd_ols <- pcd_test(
  residuals(fit_ols), data = pwt_cc,
  unit_index = "country", time_index = "year",
  test = "pesaran"
)
print(cd_ols)
```

The CD statistic is far above the critical value ($\approx 1.96$), decisively
rejecting the null of no cross-sectional dependence. Ignoring this leads to
spurious inference.

# The Model: Heterogeneous Panels with Common Factors

## The general DGP

`dcce` works with the **heterogeneous panel model** with unobserved common factors:

$$
y_{it} = \alpha_i + \boldsymbol{\beta}_i' \mathbf{x}_{it} + \boldsymbol{\lambda}_i' \mathbf{f}_t + u_{it},
\quad i = 1,\ldots,N, \quad t = 1,\ldots,T,
$$

where:

- $\boldsymbol{\beta}_i$ are **unit-specific** slope coefficients (heterogeneous);
- $\mathbf{f}_t$ is a $(m \times 1)$ vector of **unobserved common factors**;
- $\boldsymbol{\lambda}_i$ are unit-specific factor loadings;
- $u_{it}$ is an idiosyncratic error with $\mathbb{E}[u_{it} u_{jt}] = 0$ for $i \neq j$.

The regressors follow:
$$
\mathbf{x}_{it} = \mathbf{A}_i' \mathbf{g}_t + \boldsymbol{\Gamma}_i' \mathbf{f}_t + \mathbf{v}_{it},
$$
so the common factors $\mathbf{f}_t$ drive both $y_{it}$ and $\mathbf{x}_{it}$,
inducing CSD and endogeneity.

## The CCE approach

@pesaran2006estimation shows that the unobserved $\mathbf{f}_t$ can be
asymptotically spanned by the cross-sectional averages:

$$
\bar{y}_t = N^{-1}\sum_i y_{it}, \quad
\bar{\mathbf{x}}_t = N^{-1}\sum_i \mathbf{x}_{it}.
$$

These **CSAs** are observable proxies for the common factors. Augmenting each
unit's regression with the CSAs (and their lags, for dynamic panels) removes the
common factor component and restores consistency.

## Dynamic extension (DCCE)

For dynamic panels where $y_{i,t-1}$ appears as a regressor, @chudik2015common
show that the CCE estimator requires CSA lags to remain consistent. The recommended
number of lags is $p_T = \lfloor T^{1/3} \rfloor$ (roughly 3 for $T \approx 48$).
This gives the **DCCE** estimator.

# Estimators

## Mean Group (MG)

The **Mean Group** estimator of @pesaran1995estimating runs OLS separately for each
unit $i$:

$$
\hat{\boldsymbol{\beta}}_i^{\text{OLS}} = (\mathbf{X}_i' \mathbf{X}_i)^{-1} \mathbf{X}_i' \mathbf{y}_i,
$$

and averages:

$$
\hat{\boldsymbol{\beta}}^{\text{MG}} = N^{-1} \sum_{i=1}^N \hat{\boldsymbol{\beta}}_i.
$$

MG is consistent for $N, T \to \infty$ under heterogeneous slopes, but inconsistent
when CSD is present via common factors. It serves as a useful benchmark.

```{r mg-pwt8}
fit_mg <- dcce(
  data            = pwt8,
  unit_index      = "country",
  time_index      = "year",
  formula         = d_log_rgdpo ~ L(log_rgdpo, 1) + log_hc + log_ck + log_ngd,
  model           = "mg",
  cross_section_vars = NULL
)
print(fit_mg)
```

The `L(log_rgdpo, 1)` term is the first lag of log GDP---its MG coefficient is
$({\hat\rho} - 1)$, so a negative value indicates mean-reverting dynamics.

## CCE Mean Group (CCE-MG)

The **CCE-MG** estimator augments each unit's regression with contemporaneous CSAs:

$$
y_{it} = \alpha_i + \boldsymbol{\beta}_i' \mathbf{x}_{it}
          + \boldsymbol{\delta}_i' \bar{\mathbf{z}}_t + u_{it},
$$

where $\bar{\mathbf{z}}_t = (\bar{y}_t, \bar{\mathbf{x}}_t')'$. The coefficients
on the CSAs absorb the unobserved factors; the structural parameters
$\hat{\boldsymbol{\beta}}_i$ are then averaged over units.

@pesaran2006estimation establishes consistency and asymptotic normality of the
CCE-MG estimator under mild conditions.

```{r cce-pwt8}
fit_cce <- dcce(
  data               = pwt8,
  unit_index         = "country",
  time_index         = "year",
  formula            = d_log_rgdpo ~ L(log_rgdpo, 1) + log_hc + log_ck + log_ngd,
  model              = "cce",
  cross_section_vars = ~ log_rgdpo + log_hc + log_ck + log_ngd,
  cross_section_lags = 0
)
print(fit_cce)
```

Comparing MG and CCE-MG shows the effect of correcting for CSD: the coefficients
shift, reflecting the removal of common-factor-driven bias.

## Dynamic CCE (DCCE)

When the lagged dependent variable is included, @chudik2015common require CSA lags
to achieve asymptotically unbiased estimates. The DCCE estimator augments each
unit's regression with $\bar{\mathbf{z}}_t, \bar{\mathbf{z}}_{t-1}, \ldots,
\bar{\mathbf{z}}_{t-p_T}$.

```{r dcce-pwt8}
fit_dcce <- dcce(
  data               = pwt8,
  unit_index         = "country",
  time_index         = "year",
  formula            = d_log_rgdpo ~ L(log_rgdpo, 1) + log_hc + log_ck + log_ngd,
  model              = "dcce",
  cross_section_vars = ~ log_rgdpo + log_hc + log_ck + log_ngd,
  cross_section_lags = 3
)
print(fit_dcce)
```

The coefficient on `L(log_rgdpo, 1)` is the speed of adjustment in the error
correction framework. A value of approximately $-0.6$ implies $\hat\rho \approx 0.4$,
consistent with mean-reverting dynamics in log GDP.

### Verify that DCCE removes CSD

A key diagnostic is to apply the CD test to the DCCE residuals. If the estimator
has successfully absorbed the common factors, the residuals should no longer
exhibit significant CSD:

```{r dcce-cd-test}
cd_dcce <- pcd_test(fit_dcce, test = "pesaran")
print(cd_dcce)
```

After DCCE correction with three CSA lags, the CD statistic falls to an
insignificant level, confirming that the common factors have been filtered out.

## Regularized CCE (rCCE)

The **regularized CCE** of @juodis2022regularized addresses the case where the
number of CSA variables is large relative to $T$. It replaces the full CSA matrix
with its first few principal components, regularizing the factor proxy:

```{r rcce-pwt8}
fit_rcce <- dcce(
  data               = pwt8,
  unit_index         = "country",
  time_index         = "year",
  formula            = d_log_rgdpo ~ L(log_rgdpo, 1) + log_hc + log_ck + log_ngd,
  model              = "rcce",
  cross_section_vars = ~ log_rgdpo + log_hc + log_ck + log_ngd,
  cross_section_lags = 0
)
print(fit_rcce)
```

## Long-Run Estimators: CS-DL, CS-ARDL, and PMG

For researchers interested in long-run relationships, `dcce` implements three
estimators. All three produce a three-block print output: **short-run**
coefficients, the **adjustment** (speed of adjustment to equilibrium), and
**long-run** coefficients.

### CS-DL (Cross-Sectionally Augmented Distributed Lag)

The **CS-DL** estimator of @chudik2016long uses \(\Delta y_{it}\) as the
dependent variable and regresses it on the **level** of \(x_{it}\) (whose
coefficient is the long-run effect) plus the contemporaneous and lagged
first differences of \(x\) (as short-run controls) plus CSAs:

$$
\Delta y_{it} = \alpha_i + w'_i x_{it}
                + \sum_{\ell=0}^{p_x} \phi'_{i\ell} \Delta x_{i,t-\ell}
                + \delta'_i \bar{\mathbf{z}}_t + u_{it}.
$$

`dcce` **automatically** constructs \(\Delta y\) as the LHS and adds
\(\Delta x, \Delta x_{-1}, \ldots, \Delta x_{-p_x}\) when `model = "csdl"`.
You just provide the levels formula and set `csdl_xlags`:

```{r csdl-pwt8}
fit_csdl <- dcce(
  data               = pwt8,
  unit_index         = "country",
  time_index         = "year",
  formula            = log_rgdpo ~ log_ck + log_hc + log_ngd,
  model              = "csdl",
  cross_section_vars = ~ log_rgdpo + log_hc + log_ck + log_ngd,
  cross_section_lags = 3,
  csdl_xlags         = 3
)
print(fit_csdl)
```

The coefficients on `log_ck`, `log_hc`, and `log_ngd` in the **Long-run
Estimates** block are the long-run elasticities of GDP.

### CS-ARDL (Cross-Sectionally Augmented ARDL)

The **CS-ARDL** estimator fits an ARDL\((p_y, p_x)\) model with CSAs per unit
and recovers long-run coefficients and the speed of adjustment via the delta
method:

$$
y_{it} = \alpha_i + \sum_{p=1}^{P_y} \phi_{ip} y_{i,t-p}
                  + \sum_{q=0}^{P_x} \beta'_{iq} x_{i,t-q}
                  + \delta'_i \bar{\mathbf{z}}_t + e_{it},
$$

with long-run coefficients
\(\theta_{ik} = \frac{\sum_q \beta_{ikq}}{1 - \sum_p \phi_{ip}}\) and
adjustment speed \(\varphi_i = -(1 - \sum_p \phi_{ip})\). Users specify the
dynamic structure explicitly using `L()` in the formula:

```{r csardl-pwt8}
fit_csardl <- dcce(
  data               = pwt8,
  unit_index         = "country",
  time_index         = "year",
  formula            = log_rgdpo ~ L(log_rgdpo, 1)
                                   + log_hc + L(log_hc, 1)
                                   + log_ck + L(log_ck, 1)
                                   + log_ngd + L(log_ngd, 1),
  model              = "csardl",
  cross_section_vars = ~ log_rgdpo + log_hc + log_ck + log_ngd,
  cross_section_lags = 3
)
print(fit_csardl)
```

The output shows three blocks:

1. **Short-run** coefficients from the ARDL(1,1) regression (unit-level OLS, MG average)
2. **Adjustment term** \(\varphi \approx -0.69\), meaning about 69% of the deviation from
   equilibrium is corrected each period
3. **Long-run** elasticities recovered via the delta method, with MG standard errors

### PMG (Pooled Mean Group)

The **PMG** estimator of @shin1999ardl imposes **common long-run coefficients**
across units while keeping adjustment and short-run dynamics heterogeneous.
`dcce` implements PMG via inverse-variance weighted pooling of the unit-level
CS-ARDL long-run estimates — a fast, consistent alternative to the Pesaran-Shin-Smith
concentrated-MLE that avoids numerical optimization.

```{r pmg-pwt8}
fit_pmg <- dcce(
  data               = pwt8,
  unit_index         = "country",
  time_index         = "year",
  formula            = log_rgdpo ~ L(log_rgdpo, 1)
                                   + log_hc + L(log_hc, 1)
                                   + log_ck + L(log_ck, 1)
                                   + log_ngd + L(log_ngd, 1),
  model              = "pmg",
  cross_section_vars = ~ log_rgdpo + log_hc + log_ck + log_ngd,
  cross_section_lags = 3
)
print(fit_pmg)
```

Compared to CS-ARDL, PMG produces **tighter** standard errors on the long-run
coefficients because it exploits the homogeneity restriction.

## Comparing All Estimators

```{r compare-estimators}
models <- list(
  MG      = fit_mg,
  CCE     = fit_cce,
  DCCE    = fit_dcce,
  CS_DL   = fit_csdl,
  CS_ARDL = fit_csardl
)

# Extract coefficient on physical capital (log_ck) across models
key_var <- "log_ck"
compare <- data.frame(
  estimator = names(models),
  estimate  = vapply(models, function(m) {
    b <- coef(m)
    if (key_var %in% names(b)) unname(b[key_var]) else NA_real_
  }, numeric(1)),
  std_error = vapply(models, function(m) {
    se <- m$se
    if (key_var %in% names(se)) unname(se[key_var]) else NA_real_
  }, numeric(1))
)
compare$t_stat <- compare$estimate / compare$std_error
print(compare, digits = 4)
```

**Note:** MG, CCE, and DCCE estimate a **short-run** coefficient on `log_ck` within
a dynamic error-correction specification (with `L(log_rgdpo,1)` on the RHS). CS-DL
and CS-ARDL target the **long-run** elasticity of GDP with respect to capital, using
a levels-in-first-difference specification. Direct comparisons across the two groups
should be interpreted with care.

# The CD Test Suite

`dcce` implements four CD tests, each addressing different aspects of the CSD
problem.

| Test | Reference | Description |
|------|-----------|-------------|
| `pesaran` | @pesaran2015testing | Benchmark; N(0,1) under H0 |
| `cdw`     | @juodis2022 | Rademacher-weighted; robust to incidental parameters |
| `cdwplus` | Baltagi, Feng & Kao (2012) | Bias-adjusted LM with random sign weighting |
| `pea`     | @fan2015power | Power-enhanced; better against sparse alternatives |
| `cdstar`  | @pesaran2021cdstar | Bias-corrected for panels with strong factors |

```{r cd-all-tests}
# Run all five CD tests on DCCE residuals
set.seed(42)
cd_all <- pcd_test(fit_dcce,
                   test   = c("pesaran", "cdw", "cdwplus", "pea", "cdstar"),
                   n_reps = 199)
print(cd_all)
```

**Interpretation:** The four tests have different power properties:

- **Pesaran CD** is the workhorse benchmark. Insignificance ($|CD| < 1.96$) indicates
  that the estimator has removed the dominant common factor(s).
- **CDw** (Juodis-Reese) corrects for incidental parameter bias in dynamic panels;
  a value near zero confirms no systematic residual dependence.
- **PEA** is power-enhanced against **sparse** alternatives (a few strongly correlated
  pairs). It can remain significant even when the average pairwise correlation is
  near zero, so significance here need not contradict the Pesaran CD result.
- **CD\*** applies a PCA-based bias correction for panels with strong factors;
  borderline significance reflects the strength of the factor structure in the data.

The key diagnostic is the **Pesaran CD**: after DCCE with three lags, a value
close to zero (well below $\pm 1.96$) confirms that the common factors have been
successfully absorbed.

# Bootstrap Inference

Asymptotic standard errors for DCCE rest on the central limit theorem approximation
$\hat{\boldsymbol{\beta}}^{\text{MG}} \approx \mathcal{N}(\boldsymbol{\beta}, V/N)$.
In panels with small $N$, bootstrap confidence intervals are preferable.

`dcce` provides a `bootstrap()` function with two schemes:

- **Cross-section bootstrap** (default): resamples units with replacement, preserving
  time-series dynamics within each unit.
- **Wild bootstrap**: multiplies residuals by Rademacher or Mammen weights to
  preserve heteroskedasticity.

```{r bootstrap, eval=FALSE}
# Cross-section bootstrap (not evaluated in vignette to limit build time)
set.seed(42)
boot_res <- bootstrap(fit_dcce, type = "crosssection", reps = 499)
print(boot_res)
```

For a quick illustration on the simulated dataset (smaller $N$ and $T$):

```{r bootstrap-sim}
data(dcce_sim, package = "dcce")

fit_sim <- dcce(
  data               = dcce_sim,
  unit_index         = "unit",
  time_index         = "time",
  formula            = y ~ L(y, 1) + x,
  model              = "dcce",
  cross_section_vars = ~ y + x,
  cross_section_lags = 3
)

set.seed(42)
boot_sim <- bootstrap(fit_sim, type = "crosssection", reps = 199)
print(boot_sim)
```

# Worked Example: Ditzen (2018) Growth Regression

We replicate the core results from @ditzen2018estimating, Section 7, using
the `pwt8` dataset (Penn World Tables 8, N = 93 countries, T = 48 years,
1960--2007). The model is a Solow-type growth regression:

$$
\Delta \log y_{it} = \alpha_i
  + \phi_i \log y_{i,t-1}
  + \beta_{1i} \log h_{it}
  + \beta_{2i} \log k_{it}
  + \beta_{3i} \log(n_{it} + g + \delta)
  + \text{CSD correction} + u_{it},
$$

where $y$ is real GDP (output-side), $h$ is the human capital index, $k$ is the
physical capital stock, $n+g+\delta$ aggregates population growth and
depreciation, and $\phi_i = \rho_i - 1$ is the speed of adjustment.

## Step 1: Detect CSD in naive residuals

```{r ex1-ols-cd}
dat <- na.omit(pwt8[, c("country", "year", "d_log_rgdpo",
                          "log_rgdpo", "log_hc", "log_ck", "log_ngd")])
fit_ols2 <- lm(d_log_rgdpo ~ log_hc + log_ck + log_ngd + log_rgdpo, data = dat)
cd_naive <- pcd_test(residuals(fit_ols2), data = dat,
                     unit_index = "country", time_index = "year",
                     test = "pesaran")
cat(sprintf("CD statistic (OLS): %.2f  p-value: %.4f\n",
            cd_naive$statistics$statistic[1],
            cd_naive$statistics$p_value[1]))
```

**Result:** The CD statistic is large and significant, confirming strong CSD.

## Step 2: Mean Group (no CSD correction)

```{r ex2-mg}
fit_mg2 <- dcce(
  data               = pwt8,
  unit_index         = "country",
  time_index         = "year",
  formula            = d_log_rgdpo ~ L(log_rgdpo, 1) + log_hc + log_ck + log_ngd,
  model              = "mg",
  cross_section_vars = NULL
)

cd_mg <- pcd_test(fit_mg2, test = "pesaran")
cat(sprintf("MG: phi = %.3f, log_ck = %.3f\n",
            coef(fit_mg2)["L(log_rgdpo,1)"], coef(fit_mg2)["log_ck"]))
cat(sprintf("CD after MG: %.2f  p-value: %.4f\n",
            cd_mg$statistics$statistic[1], cd_mg$statistics$p_value[1]))
```

MG residuals still exhibit significant CSD because MG does not correct for common
factors.

## Step 3: Dynamic CCE with 3 CSA lags (Ditzen 2018 Example 7.3)

```{r ex3-dcce}
fit_dcce2 <- dcce(
  data               = pwt8,
  unit_index         = "country",
  time_index         = "year",
  formula            = d_log_rgdpo ~ L(log_rgdpo, 1) + log_hc + log_ck + log_ngd,
  model              = "dcce",
  cross_section_vars = ~ log_rgdpo + log_hc + log_ck + log_ngd,
  cross_section_lags = 3
)

cd_dcce2 <- pcd_test(fit_dcce2, test = "pesaran")
cat(sprintf("DCCE: phi = %.3f, log_ck = %.3f\n",
            coef(fit_dcce2)["L(log_rgdpo,1)"], coef(fit_dcce2)["log_ck"]))
cat(sprintf("CD after DCCE(3): %.2f  p-value: %.4f\n",
            cd_dcce2$statistics$statistic[1], cd_dcce2$statistics$p_value[1]))
```

After applying DCCE with three CSA lags, the CD statistic is no longer
significant, indicating that the estimator has successfully absorbed the
common factors.

## Step 4: Summary table

```{r ex4-summary-table}
# Coefficient table comparing MG and DCCE
vars <- c("L(log_rgdpo,1)", "log_hc", "log_ck", "log_ngd")
tab <- data.frame(
  Variable = vars,
  MG_coef  = round(coef(fit_mg2)[vars], 3),
  MG_se    = round(fit_mg2$se[vars], 3),
  DCCE_coef = round(coef(fit_dcce2)[vars], 3),
  DCCE_se   = round(fit_dcce2$se[vars], 3),
  row.names = NULL
)
tab$MG_sig   <- ifelse(abs(coef(fit_mg2)[vars] / fit_mg2$se[vars]) > 1.96, "*", "")
tab$DCCE_sig <- ifelse(abs(coef(fit_dcce2)[vars] / fit_dcce2$se[vars]) > 1.96, "*", "")
print(tab)
```

**Interpretation:**

- The **speed of adjustment** ($\hat\phi$) is negative and significant, confirming
  mean-reverting dynamics in log GDP. The DCCE estimate ($\approx -0.6$) implies
  $\hat\rho \approx 0.4$, plausible for this sample.
- **Physical capital** (`log_ck`) has a positive, significant effect in both
  specifications, consistent with the Solow model.
- **Population growth** (`log_ngd`) is negative, also in line with the Solow model.
- The shift in coefficient magnitudes from MG to DCCE reflects the bias from
  CSD in the MG estimator.

## Step 5: Unit-level heterogeneity

One advantage of the Mean Group approach is access to unit-level estimates.
We can examine the distribution of the speed-of-adjustment across countries:

```{r ex5-unit-coefs, fig.alt="Histogram of unit-level speed of adjustment coefficients"}
b_unit <- coef(fit_dcce2, type = "unit")
ar_coefs <- b_unit$estimate[b_unit$term == "L(log_rgdpo,1)"]
rho_implied <- ar_coefs + 1  # rho = (phi + 1)

cat(sprintf("Implied rho: mean = %.3f, median = %.3f, [min, max] = [%.3f, %.3f]\n",
            mean(rho_implied), median(rho_implied),
            min(rho_implied), max(rho_implied)))
cat(sprintf("Fraction with 0 < rho < 1 (stationarity): %.1f%%\n",
            100 * mean(rho_implied > 0 & rho_implied < 1)))

hist(
  rho_implied,
  breaks = 20,
  main   = "Unit-level AR(1) coefficient (rho = phi + 1)",
  xlab   = expression(hat(rho)[i]),
  col    = "steelblue",
  border = "white"
)
abline(v = median(rho_implied), col = "firebrick", lwd = 2, lty = 2)
legend("topright", legend = "Median", col = "firebrick", lty = 2, lwd = 2, bty = "n")
```

The distribution is concentrated in $(0, 1)$, confirming stationary dynamics for
most countries, with the median $\hat\rho \approx 0.4$.

# Auxiliary Diagnostics

## Exponent of cross-sectional dependence

The **CSD exponent** $\alpha$ of @bailey2016exponent characterizes the strength of
dependence. For $\alpha = 1$ the panel exhibits strong CSD; for $\alpha < 1/2$ the
CSD is weak and standard estimators remain consistent.

```{r csd-exp}
# CSD exponent of log real GDP (levels) across countries
data(pwt8)
X_lev <- do.call(rbind, by(pwt8, pwt8$country, function(d) d$log_rgdpo))
# Keep only time periods with no missing values
X_lev <- X_lev[, colSums(is.na(X_lev)) == 0, drop = FALSE]
alpha_res <- csd_exp(X_lev, use_residuals = FALSE)
print(alpha_res)
```

An estimate of $\hat\alpha \approx 0.84$ is close to 1, signalling strong CSD in log
real GDP---further motivation for using DCCE. For first-differenced data (growth
rates) the exponent is typically much lower because differencing strips out the
common trend.

## Information criterion for CSA lag selection

@margaritella2023information propose IC and PC criteria for selecting the number of
CSA variables in **static** CCE models. We compare static CCE models with 0--3
contemporaneous CSA variables (via varying the number of regressors included in
`cross_section_vars`). For **dynamic** panels, the Chudik-Pesaran rule
$p_T = \lfloor T^{1/3} \rfloor$ is the standard choice for CSA lags.

```{r ic-selection}
# Compare static CCE at lag 0 across different CSA variable sets
# (IC criteria apply to the static CCE case)
lags_to_try <- 0:3
models_ic <- lapply(lags_to_try, function(p) {
  dcce(
    data               = pwt8,
    unit_index         = "country",
    time_index         = "year",
    formula            = d_log_rgdpo ~ L(log_rgdpo, 1) + log_hc + log_ck + log_ngd,
    model              = "cce",
    cross_section_vars = ~ log_rgdpo + log_hc + log_ck + log_ngd,
    cross_section_lags = p
  )
})

ic_values <- sapply(models_ic, function(m) {
  res <- ic(m, models = models_ic)
  c(IC1 = res$IC1, IC2 = res$IC2)
})
ic_table <- data.frame(lags = lags_to_try, t(ic_values))
print(ic_table, digits = 5)
cat(sprintf("IC1 selects %d CSA lag(s)\n", lags_to_try[which.min(ic_table$IC1)]))
cat(sprintf("IC2 selects %d CSA lag(s)\n", lags_to_try[which.min(ic_table$IC2)]))
```

A value of 0 selected by IC indicates that contemporaneous CSAs suffice for the
static CCE specification---consistent with the Pesaran (2006) theoretical
recommendation. For the dynamic DCCE application above, we use 3 lags as prescribed
by the Chudik-Pesaran $p_T = \lfloor T^{1/3} \rfloor \approx 3$ rule.

## Rank condition

The **rank condition** of @devos2024rank checks whether the CSA variables span the
factor space, a necessary condition for CCE consistency:

```{r rank-condition}
rc <- rank_condition(fit_cce)
cat(sprintf("Estimated factors (m): %d\n", rc$m))
cat(sprintf("Rank of avg loadings (g): %d\n", rc$g))
cat(sprintf("Rank condition (RC = 1 means satisfied): %d\n", rc$RC))
```

## Panel unit root test: Pesaran CIPS

The **CIPS** test of @pesaran2007simple is the workhorse panel unit root test
under cross-sectional dependence. It augments unit-level Dickey-Fuller
regressions with cross-sectional averages and averages the resulting
CADF t-statistics.

```{r cips-test}
X_lev <- do.call(rbind, by(pwt8, pwt8$country, function(d) d$log_rgdpo))
X_lev <- X_lev[, colSums(is.na(X_lev)) == 0, drop = FALSE]
cips_res <- cips_test(X_lev, lags = 1, trend = FALSE)
print(cips_res)
```

A CIPS statistic below the 5% critical value of \(-2.27\) rejects the null of a
panel unit root.

## Slope heterogeneity: Swamy and Pesaran-Yamagata

Before deciding between MG and a pooled estimator, you should formally test
whether slopes are homogeneous. The **Swamy (1970)** test and its standardised
large-\(N\) version due to **Pesaran & Yamagata (2008)** are implemented as
`swamy_test()`:

```{r swamy-test}
swamy_test(fit_cce)
```

A large, significant statistic rejects homogeneity and justifies heterogeneous
slopes (MG/CCE/DCCE). For this growth regression the test rejects decisively,
confirming that growth dynamics differ across countries.

## Hausman test: MG vs pooled

A Hausman-style comparison of MG against the weighted pooled estimator:

```{r hausman-test}
hausman_test(fit_cce)
```

Under the null (homogeneous slopes) both estimators are consistent; under the
alternative only MG is. Rejection of the null is another signal that the
homogeneous-slope assumption is too restrictive.

## `broom` compatibility

`dcce_fit` objects are compatible with `broom::tidy()` and `broom::glance()`.
For long-run estimators, `tidy()` additionally includes rows for the adjustment
speed and the long-run coefficients, tagged in the `type` column:

```{r broom-methods}
tidy(fit_dcce2)
glance(fit_dcce2)
```

```{r broom-csardl}
# For a CS-ARDL fit, tidy() returns short-run, adjustment, and long-run rows
tidy(fit_csardl)
```

## Confidence intervals

The `confint()` method supports three types: `"mg"` for the main coefficients,
`"lr"` for long-run estimates, and `"adjustment"` for the speed of adjustment:

```{r confint-demo}
confint(fit_csardl, type = "lr", level = 0.90)
confint(fit_csardl, type = "adjustment")
```

## Visualising heterogeneity

The `plot()` method shows the distribution of unit-level coefficients:

```{r plot-coef, fig.alt="Histograms of unit-level coefficients for the DCCE growth regression"}
plot(fit_dcce2, which = "coef")
```

Each panel corresponds to one regressor; the dashed red line marks the MG mean.
Use `which = "resid"` for a residual diagnostic plot.

# Additional Estimators (v0.2.0+)

## Augmented Mean Group (AMG)

The **AMG** estimator of @eberhardt2010productivity accounts for CSD by
extracting a Common Dynamic Process (CDP) from time-dummy coefficients in a
pooled first-difference regression. The CDP is then added as a nuisance
regressor in each unit's OLS:

```{r amg}
fit_amg <- dcce(
  data = pwt8, unit_index = "country", time_index = "year",
  formula = d_log_rgdpo ~ log_hc + log_ck + log_ngd,
  model = "amg", cross_section_vars = NULL
)
coef(fit_amg)
```

## Interactive Fixed Effects (IFE, Bai 2009)

Unlike CCE (which proxies factors via cross-sectional averages), **IFE**
estimates the factors and loadings directly via iterative principal
components. It produces **pooled** slope coefficients (not heterogeneous):

```{r ife}
fit_ife <- dcce(
  data = pwt8, unit_index = "country", time_index = "year",
  formula = log_rgdpo ~ log_hc + log_ck + log_ngd,
  model = "ife", n_factors = 2L
)
print(fit_ife)
```

## Pooled CCE (CCEP)

@pesaran2006estimation proposed two CCE estimators: the Mean Group version
(CCE-MG, our `model = "cce"`) and the **Pooled** version (CCEP). CCEP
constrains slopes to be identical across units, gaining efficiency when
slopes are truly homogeneous:

```{r ccep}
fit_ccep <- dcce(
  data = pwt8, unit_index = "country", time_index = "year",
  formula = d_log_rgdpo ~ log_hc + log_ck + log_ngd,
  model = "ccep", cross_section_vars = ~ .
)
coef(fit_ccep)
```

# Additional Diagnostics (v0.3.0+)

## Panel Granger causality

The Dumitrescu-Hurlin (2012) test checks whether one variable
Granger-causes another in a heterogeneous panel:

```{r granger}
gc <- granger_test(
  data = pwt8, unit_index = "country", time_index = "year",
  y = "d_log_rgdpo", x = "log_ck", lags = 1L
)
print(gc)
```

## IPS and LLC panel unit root tests

Standard (non-CSD-robust) benchmarks alongside CIPS:

```{r ips-llc}
X_lev <- do.call(rbind, by(pwt8, pwt8$country, function(d) d$log_rgdpo))
X_lev <- X_lev[, colSums(is.na(X_lev)) == 0, drop = FALSE]
panel_ur_test(X_lev, test = "ips", lags = 1)
```

## Pedroni cointegration test

```{r pedroni}
panel_coint_test(
  data = pwt8, unit_index = "country", time_index = "year",
  formula = log_rgdpo ~ log_hc + log_ck,
  test = "pedroni", lags = 1
)
```

## Structural break test

The sup-Wald test for an unknown break date with Andrews (1993) critical
values:

```{r structural-break}
brk <- structural_break_test(
  data = pwt8, unit_index = "country", time_index = "year",
  formula = d_log_rgdpo ~ log_hc + log_ck + log_ngd,
  model = "mg", cross_section_vars = NULL,
  type = "unknown", trim = 0.20
)
print(brk)
```

# Impulse Response Functions (v0.4.0)

For dynamic models, `irf()` traces the response of the dependent variable
to a unit shock in a regressor:

```{r irf-example}
data(dcce_sim)
fit_dyn <- dcce(
  data = dcce_sim, unit_index = "unit", time_index = "time",
  formula = y ~ L(y, 1) + x,
  model = "dcce", cross_section_vars = ~ ., cross_section_lags = 3
)
ir <- irf(fit_dyn, impulse = "x", horizon = 10, boot_reps = 99, seed = 42)
print(ir)
```

# The `dcce_workflow()` Pipeline

A single function runs all recommended pre-estimation diagnostics and
returns a recommended `dcce()` call:

```{r workflow-demo, eval=FALSE}
dcce_workflow(
  data = pwt8, unit_index = "country", time_index = "year",
  formula = log_rgdpo ~ log_hc + log_ck + log_ngd
)
```

# Quick-Start Guide

```{r quickstart, eval=FALSE}
library(dcce)
data(pwt8)

# 1. Detect CSD
fit_ols_qs <- lm(d_log_rgdpo ~ log_hc + log_ck + log_ngd, data = na.omit(pwt8))
pcd_test(residuals(fit_ols_qs), data = na.omit(pwt8),
         unit_index = "country", time_index = "year",
         test = "pesaran")

# 2. Estimate with DCCE
fit_qs <- dcce(
  data               = pwt8,
  unit_index         = "country",
  time_index         = "year",
  formula            = d_log_rgdpo ~ L(log_rgdpo, 1) + log_hc + log_ck + log_ngd,
  model              = "dcce",
  cross_section_vars = ~ log_rgdpo + log_hc + log_ck + log_ngd,
  cross_section_lags = 3
)

# 3. Check residuals
pcd_test(fit_qs, test = "pesaran")

# 4. Tidy output
tidy(fit_qs)
glance(fit_qs)

# 5. Bootstrap
dcce_bootstrap(fit_qs, reps = 499)
```

# References

<div id="refs"></div>
